---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# pkgwatch

<!-- badges: start -->
[![Travis build status](https://travis-ci.com/coatless/pkgwatch.svg?branch=master)](https://travis-ci.com/coatless/pkgwatch)
[![CRAN status](https://www.r-pkg.org/badges/version/pkgwatch)](https://CRAN.R-project.org/package=pkgwatch)
[![Codecov test coverage](https://codecov.io/gh/coatless/pkgwatch/branch/master/graph/badge.svg)](https://codecov.io/gh/coatless/pkgwatch?branch=master)
<!-- badges: end -->

The goal of pkgwatch is to prevent _R_ packages from being used through
`library()` and `require()` calls. This is useful to prevent (un)intentional
use of packages not approved within a classroom, research, or company environment.

## Installation

<!--
You can install the released version of pkgwatch from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("pkgwatch")
```

Or, you can be on the cutting-edge development version on GitHub using: -->

This package is only available on GitHub. To install the package,
please use:

```{r gh-installation, eval = FALSE}
if(!requireNamespace("remotes")) install.packages("remotes")
remotes::install_github("coatless/pkgwatch")
```

## Usage

To use the `pkgwatch` package, load it into _R_ using:

```{r example}
library("pkgwatch")
```

From there, any package that is on a blacklist will be prevented from
being loaded. The blacklist can be established on a per-session basis or 
can be loaded as needed. 

For example, let's say we didn't want to allow `toad` to be loaded. We would
call:

```{r add-pkg}
watch_add("toad")
```

If we attempted to load `toad` using either `library()` or `require()`, then
we would error:

```r
library("toad")
#> Detected {toad} package load...
#> The {toad} package is not allowed to be used.
#> Error in as.environment(lib.pos) : invalid 'pos' argument

require("toad")
#> Loading required package: toad
#> Detected {toad} package load...
#> The {toad} package is not allowed to be used.
#> Failed with error:  'invalid 'pos' argument'
```

All packages that are prohibited from being used can be viewed with:

```{r list-watches}
watch_active()
```

To allow the package to be used, we would need to remove the watch:

```{r remove-watch}
watch_remove("toad")
```

Then, the package load would be allowed. 

```r
library("toad")
```

## Motivation

When designing `pkgwatch`, the goal was to achieve a "soft-failure" when 
undesirable packages were loaded via `library()` or `require()`. 
Generally, this follows in the footsteps of 
[`strict`](https://github.com/hadley/strict)
-- which sought to raise issues with undesirable design patterns in code --
and, subsequently, [`conflicted`](https://github.com/r-lib/conflicted) --
which addressed search path collisions between similarly named functions in
different packages -- both by Hadley Wickham. With this being said,
there are better variants of protecting the _R_ process. Most notably, the
[`RAppArmor`](https://cran.r-project.org/package=RAppArmor) by Jeroen Ooms
provides superior sandboxing of _R_. Alternatively, the version of _R_ 
could simply not have these packages installed to begin with.

Fun fact: The code for this sat in an `untitled.R` file for ~2 years.

## Author

James Joseph Balamuta

## License

GPL (>= 2)
